---
title: "Course Project"
subtitle: "Developing Data Products"
author: "Malu Mondelli"
date: "October 11, 2017"
output: ioslides_presentation
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

- This presentation is being created as part of the final project of the Developing Data Products course (Coursera).
- The idea is to present a shiny application developed for this project as well as its main characteristics.
- The application presents graphically and interactively some statistics about the internet service in Brazil.

### About the Dataset

- The dataset was obtained through the 'transparency portal' maintained by the Brazilian government and contains data on the internet service offered by different companies.
- Available at: http://dados.gov.br/dataset/resultado-consolidado-da-pesquisa-banda-larga-fixa


## The application

### The main features of the application are:

- The Statistics menu, where the user can choose an internet company to analyze the percentage of service channels used in each state by year (2015 and 2016).
- By choosing a state, the user can view the percentage of scores on the quality of service, assigned by users in that state (for a specific company).
- Also, the application allows access and filtering the dataset according to the user's interest in the Dataset menu.

## Example of statistic extracted from the dataset

- A calculation allows to obtain the coverage of each company (number of states where each company offers the service):

```{r, eval=TRUE}
blf.df <- read.csv('./Data/banda_larga_fixa2.csv', 
               header=T, sep=';',
               fileEncoding = "UTF-8"); topic <- unique(blf.df$TEMA)
aux <- unique(blf.df[blf.df$TEMA == topic[1],])
calc <- aux[,c(2,3)]; calc <- calc[!grepl("Total", calc$ESTADO),]
calc <- unique(calc)
calc <- aggregate(ESTADO ~ OPERADORA, data = calc, FUN = function(x){NROW(x)})
colnames(calc) <- c('Company', 'States'); head(calc)

```


## Useful Links

- Source Code: https://github.com/mmondelli/datasciencecoursera/tree/master/Data_Products/Project
- This presentation: https://mmondelli.github.io/datasciencecoursera/Data_Products/Project/presentation.html
- The application: https://mmondelli.shinyapps.io/data_products_project_-_coursera/
- Any questions, feel free to contact me: mluiza@lncc.br

### Thanks!
